---
title: 'Econ 590 (M2): Assignment 7'
author: "Zhixin Lin"
date: "October 20, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

* Total points: 100
* Due Date: October 26, 2018 at 10:05am
* Hand  in a .html copy on Compass 2G OR a hard copy in class
* Be succinct and concise in your answers (more words is not always better!).


# Question 1: Cognitive Behavioral Therapy and Jail TIme (analysis) (100 points total)

In recent years, Cognitive Behavioral Therapy (CBT) has become an increasingly discussed way to reduce criminal activity amongst young people around the world.  CBT is a therapy that "focuses on challenging and changing unhelpful thoughts, beliefs, attitudes, and behaviors, and aims to develop personal coping strategies that target solving current problems."  The hope is that CBT can improve emotioanl regulation and decisionmaking, reducing the impulsive decisions that so often lead to crimial activity. You receive data from young adults in Champaign on whether or not they receuved CBT in 2016 and, subsequently, whether they spent anytime in jail in 2017.  You also have data on their age, gender, family income, and on their number of previous arrests before 2016 (i.e. 2015 and before).  Note that these data do NOT come from a randomized controlled trial.

*Type your answers below using markdown*


```{r, results = 'asis',echo=FALSE}
rmarkdownTable <- function(df){
  cat(paste(names(df), collapse = "|"))
  cat("\n")
  cat(paste(rep("-", ncol(df)), collapse = "|"))
  cat("\n")

  for(i in 1:nrow(df)){
    cat(paste(df[i,], collapse = "|"))
    cat("\n")
    }
invisible(NULL)
}

id <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16) # young person id #
y1 <- c(1,0,0,0,1,0,0,0,0,0,0,1,1,0,0,1) # indicator for going to jail when receiving CBT
y0 <- c(1,0,0,1,1,1,1,0,1,0,0,0,0,1,1,1) # indicator for going to  jail when NOT receiving CBT
D <- c(1,0,0,0,0,1,1,0,0,1,1,0,0,1,1,1) # D=1, receives CBT, D=0, does not receive CBT
female <- c(0,1,1,0,1,0,0,0,0,1,1,1,1,0,0,0) #1 = female, 0=male 
age <- c(2,3,3,3,2,3,1,1,1,2,1,3,2,3,1,3) # Age: 1 = less than 18, 2 = 18-25, 3=25+
previous.arrests <- c(12,0,0,5,7,4,8,0,0,0,0,0,0,7,14,6) # total number of arrests before receiving CBT
family.income <- c(45,10,60,100,20,0,30,45,20,10,80,30,90,40,20,25) # family income (1000s)
observed = c(1,0,0,1,1,0,0,0,1,0,0,0,0,0,0,1)
potentialoutcomes <- data.frame(id,observed,y1,y0,D,female,age,previous.arrests,family.income)
colnames(potentialoutcomes) <- c("Person id #","Observed","Y(1)","Y(0)","D","Female","Age","# of Previous Arrests","Family Income")
rmarkdownTable(potentialoutcomes)



```
 
(a) (4 points) Describe potential outcomes in this setting in words.
* Y1i : Whether or not individual i in Champaign spend time in jail if he or she receuved CBT in 2016
* Y0i : Whether or not individual i in Champaign spend time in jail if he or she did not receuved CBT in 2016
  
(b) (6 points) What is the ATE in this setting?  What is the ATET?  What is the ATENT?
* ATE = E[Y1i-Y0i]: the average treatment effect of receiving CBT for the full young adults population among young adults in champaign area.
* ATET = E[Y1i-Y0i|Di=1] the average treatment effect of receiving CBT for young adults population who received CBT in 2016,champaign area
* ATENT = E[Y1i-Y0i|Di=0]the average treatment effect of receiving CBT for young adults population who did not receive CBT in 2016,champaign area.

(c) (4 points) What are observed outcomes for each person? (i.e. calculate a new vector which is the observed outcome for each person)
* Observed outcomes for each person is 1,0,0,1,1,0,0,0,1,0,0,0,0,0,0,1
(d) (6 points) You estimate a linear regression of observed outcomes on the # of previous arrests.  The coefficient you estimate is .03.  What are the fitted values for each observation?  What are the residuals?
```{r}
observed <- c(1,0,0,1,1,0,0,0,1,0,0,0,0,0,0,1)
fit1 <- lm(observed~previous.arrests)
summary(fit1)
fit1$fitted.values
fit1$residuals
```
(e) (12 points) What is a confounding variable?  What variable(s) is(are) confounding variable(s) in the table above?  Provide quantitative evidence for your claims.
* Confouding variable: varible that being omitted in the experiment, affect both treatment and potential outcome.
* Confounding variables: Age,Female,Family income, # of Previous Arrests.
* Quantitative Evidence: What shows below is the p-values for confounding variables,each confounding variables run regressoion over both treatment and potential outcomes. If one p-value(s) is significant, that means it can cause omitted variable bias.
```{r}
#this is for age confounding variable
summary(lm(observed~D+age))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on age significant or not,regress on observed outcome
summary(lm(D~ age))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on age significant or not,regress on D
```

```{r}
#this is for age confounding variable
summary(lm(observed~D+female))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on female significant or not,regress on observed outcome
summary(lm(D~ female))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on female significant or not,regress on D
```
```{r}
#this is for age confounding variable
summary(lm(observed~D+previous.arrests))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on previous.arrests significant or not,regress on observed outcome
summary(lm(D~ previous.arrests))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on previous.arrests significant or not,regress on D
```
```{r}
#this is for age confounding variable
summary(lm(observed~D+family.income))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on family.income significant or not,regress on observed outcome
summary(lm(D~ family.income))$coefficients[,'Pr(>|t|)'] #this is to show whether the coeffcients on family.income significant or not,regress on D
```

(f) (12 points) Among the confounding variables you identified in part (e), is there full overlap for each of these (individually)?  How does overlap (or the lack there of) change the estimator in each case?  What is the confounding variable for which you have the most overlap?

    + No,there exists some variables that doesn't have full overlap.
        + For Female:  Full overlap.
        + For Age: Full overlap.
        + number of Previous Arrests: Not full overlap
        + Family Income: Not full overlap, exists groups only have treated or untreated groups.
    + Age and Femalehas the most overlap.
(g) (16 points) Using the confounding variable you identified in part (f) as having the best overlap, construct an exact matching variable for the effects of cognitive behavioral therapy (D) on whether or not the individual is sent to jail. What do you estimate?  What is the standard error of this estimate?   
* In the subclassification of age 1: ATE(1)=-1/2, s.e.= 1/5
* In the subclassification of age 2: ATE(2)=0, s.e.=1/4 + 1/4 = 1/2
* In the subclassification of age 3: ATE(3)=1/12, s.e.= 2/7 + 3/7 = 5/7
* Var(ATE)= 1/8 + 5/16 + 1/16 = 1/2
* ATE = -0.276
(h) (14 points) You instead decide that you want to exact match using both  of the confounding variables  you identifies.  How many unique combinations of thesevariables are there?  Can you exact match using these variables?  Why or why not?  What challenge does this illustrate?
 + 6 combinations.
    + cannot exact match.
    + there exists some observations cannot full matched.
    + challenge: This will present we cannot calculate certain subgroups' $ATE$.


(i) (10 points) To solve the problem in part (f), you decide to estimate a propensity score model of the probability of being treated using age-group, previous arrests, gender, and family income.  You estimate logit coefficients of .3126 for previous arrests, -.28 for gender, -.43 for age, and -.01 for family income.  What is the estimated propensity score for each observation?
```{r}
logit <- glm(D~female+age+previous.arrests+family.income, family=binomial(link="logit"))
summary(logit)
pscore <- logit$fitted.values
pscore
```

(j) (10 points) Estimate the treatment effect using subclassification on the propensity score using the estimated propensity score from part (g), grouping observations into three equal propentiy score groups (where group 1 has propensity scores  0-.33, group 2: .34-.66, and group 3: .67 to 1).  What is your estimate?  
* Group1: 2,3,4,11,12,13
* Group2: 6,8,9,10
* Group3: 1,5,7,14,15,16
* ATE(1)=-1/5  ATE(2)=-1/2  ATE(3)=-3/5
* ATE = -17/40 = -0.425
(k) (8 points) Describe two ways, other than using propensity scores, that you could have used to solve the challenge discusssed in part (h)?  
* Nearest neighbor matchinng 
* Functional Form/Dimension reduction 




